### 2016-06-06 transformation.tex
edit section 3
edit section 4.1

\section{Deriving a Corresponding Deferrable Task Set and Schedulability Test}
\label{sec:convert}

The example in Section~\ref{sec:unschedulable} can be easily converted to a corresponding deferrable task set, as explained at the end of Section~\ref{sec:unschedulable}.  Due to Theorem 5 in~\cite{Raj:suspension1991}, the feasibility of the schedule by the period enforcer algorithm depends on whether the corresponding deferrable task set can be feasibly scheduled. Therefore, before applying the period enforcer algorithm to handle segmented self-suspending sporadic tasks, we need to first derive the corresponding deferrable task set and perform the schedulability test for the deferrable task set. To reduce the pessimism of the transformation to the corresponding deferrable task set, it would be the best to have a precise transformation and an exact schedulability test. Furthermore, as demonstrated with the example shown in Figure~\ref{fig:example}, even if the conversion is done precisely, the transformed single-segment deferrable task set can admit more pessimism than the original self-suspending task set with respect to schedulability.

With the above discussions, we need to answer two questions before adopting the period enforcer algorithm. First, how can we derive the corresponding deferrable task set efficiently? Second, how should we perform the schedulability test on the corresponding deferrable task set correctly? We will discuss the difficulty to derive the corresponding deferrable task set in Section~\ref{sec:transformation-exponential} and the potential pessimism for the schedulability test based on the deferrable task set in Section~\ref{sec:schedulability-test-deferrable}.

@@

The transformation from a segmented self-suspending task set to a corresponding deferrable task set was not discussed in \cite{Raj:suspension1991}. However, in our opinion, in general, such a transformation is not an easy problem. We demonstrate the inherent difficulty by focusing on a special case and by applying the recent result provided by Nelissen et al. \cite{ecrts15nelissen}, which analyzed the exact worst-case response time for segmented self-suspending sporadic tasks with exponential time complexity. The worst-case response time analysis in \cite{ecrts15nelissen} is exact under the following conditions: 1) the task set has only one segmented self-suspending task, 2) the self-suspending task is the lowest-priority task, 3) the scheduling policy is preemptive fixed-priority scheduling, and 4) $D_i \leq T_i$ for all task $\tau_i$ in the task system.\footnote{Here, we refer to the characteristics of the worst-case release pattern provided in Lemma 2 in \cite{ecrts15nelissen}. The exact worst-case response time is to explore all the release patterns that satisfy the conditions in Lemma 2 in \cite{ecrts15nelissen} instead of the proposed MILP in \cite{ecrts15nelissen} (which only provides an upper bound). }

Suppose that the system has $k-1$ regular sporadic tasks and only one segmented self-suspending task $\tau_k$. The implicit deadline of a task is equal to its minimum inter-arrival time. Suppose that task $\tau_k$ has $m_k$ segments with $m_k \geq 3$.  Converting a computation segment into a deferrable task requires deriving the worst-case deferrable time (release jitter), denoted as $\rho_k^j$, for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$. Formally, if a job of task $\tau_k$ arrives at time $t$, it is guaranteed that the $j^{\mathrm{th}}$ computation segment of this job will arrive no later than $t+\rho_k^j$. Suppose that the worst-case response time of the $j^{\mathrm{th}}$ computation segment of task $\tau_k$ is $W_k^j$. Therefore,  if we can derive the exact $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ in this special case, we can clearly conclude that $\rho_k^1=0$ and $\rho_k^j = W_k^{j-1}+S_k^{j-1}$ for $j=2,3,\ldots,m_k$.

Based on these considerations, it appears that, at least for the simple example, the problem is basically identical to the worst-case response time analysis of segmented self-suspending task systems. Deriving the exact $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ is not an easy problem.  The method recently provided by Nelissen et al. \cite{ecrts15nelissen} can be used for this specific case to derive the exact $W_k^j$ if we assume that task $\tau_k$ has only the first $j$ computation segments and $j-1$ self-suspension intervals.  However, Nelissen et al. \cite{ecrts15nelissen} also showed that calculating the worst-case response time in the above ``simple'' case is already a very challenging problem, in which calculating $W_k^j$ would need exponential time complexity if $j \geq 2$.
In particular, Nelissen et al. \cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be a very challenging problem.\footnote{The schedulability test problem for such a case has been recently proved to be coNP-hard in the strong sense by Jian-Jia Chen in a report.}

In the context of the period enforcer, we consequently observe that the only existing solution for deriving the \emph{precise} bound $W_k^{j}$ (and hence $\rho_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity (even for the special case above).
However, finding a safe upper bound of $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ can be done in pseudo-polynomial time \cite{PH:rtss98,Huang:multiseg}, if over-approximations can be tolerated.

>>>>

\section{Deriving and Analyzing a Corresponding Deferrable Task Set}
\label{sec:convert}

To apply an analysis of the period enforcer based on Theorem 5 in~\cite{Raj:suspension1991}, we need to convert a given set of multi-segment self-suspending tasks into a corresponding set of single-segment deferrable tasks. This raises two questions: first, how can we efficiently derive the corresponding set of single-segment deferrable tasks? And second, how should we perform the schedulability test on the corresponding single-segment deferrable tasks correctly? We will discuss the difficulty of deriving a suitable corresponding task set in Section~\ref{sec:transformation-exponential} and the potential for pessimism in any schedulability test based on the set of single-segment deferrable tasks  in Section~\ref{sec:schedulability-test-deferrable}.

@@

The needed transformation from a multi-segmented self-suspending task to a corresponding set of single-segment deferrable tasks was not discussed in \cite{Raj:suspension1991}. However, in our opinion, performing such a transformation without introducing additional pessimism is not easy in the general case, nor is it obvious how to do so.

We demonstrate the inherent difficulty of the problem by focusing on a special case to which we can apply a recent result of Nelissen et al. \cite{ecrts15nelissen}, which allows analyzing the exact worst-case response time of multi-segmented self-suspending sporadic tasks, albeit with exponential time complexity. 
Nelissen et al.'s worst-case response time analysis~\cite{ecrts15nelissen} is exact under the following conditions:\footnote{We refer to the characteristics of the worst-case release pattern provided in Lemma 2 in \cite{ecrts15nelissen}. The exact worst-case response time can be obtained by exploring all  release patterns that satisfy these conditions.} 
\begin{enumerate}
	\item the task set contains only one self-suspending task, 
	\item the self-suspending task is the lowest-priority task, 
	\item the scheduling policy is preemptive fixed-priority scheduling, and 
	\item all tasks have constrained deadlines.	
\end{enumerate}

Suppose that the system has $k-1$ regular sporadic tasks and only one segmented self-suspending task $\tau_k$, and that all tasks have implicit deadlines ($T_i = D_i$ for all $\tau_i$). Further suppose that task $\tau_k$ has $m_k$ segments with $m_k \geq 3$.  To  convert a computation segment into a single-segment deferrable task, we need to derive the segment's \emph{latest-possible arrival time}, relative to the release of the job. Formally,  for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$, we let $\rho_k^j$ denote its latest-possible arrival time, with the interpretation that, if a job of task $\tau_k$ arrives at time $t$, then  it is guaranteed that the $j^{\mathrm{th}}$ computation segment of this job will arrive no later than at time $t+\rho_k^j$.

How can we compute $\rho_k^j$? Suppose that the worst-case response time of the $j^{\mathrm{th}}$ computation segment of task $\tau_k$ is $W_k^j$, and recall that $S_k^{j}$ denotes the maximum self-suspension length before the $j^{\mathrm{th}}$ computation segment of $\tau_k$. Then $\rho_k^j$ can be expressed in terms of $W_k^{j-1}$:
$$
	\rho_k^j = W_k^{j-1}+S_k^{j-1},
$$
where $W_k^0$.  Therefore, if we can derive the exact $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$, we can easily compute $\rho_k^j$  for $j=1,2,\ldots,m_k$. And conversely, if we can somehow compute $\rho_k^j$  for $j=2,\ldots,m_k$, we  can trivially infer $W_k^j$ for $j=1,2,\ldots,m_k-1$.
Based on these considerations, it appears that the transformation problem is  --- at least in the considered special case --- equivalent to worst-case response time analysis of multi-segmented self-suspending task systems.

However, deriving an exact bound $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ is not easy at all: 
even for the above ``simple'' case, Nelissen et al.'s solution~\cite{ecrts15nelissen} for calculating the exact worst-case response time requires exponential time complexity if $j \geq 2$. Furthermore, Nelissen et al. \cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be a very challenging problem indeed.\footnote{In fact, in ongoing work, this problem has recently been shown to be coNP-hard in the strong sense~\cite{Chen2016b}.}

In the context of the period enforcer, we consequently observe that the only existing solution for deriving the \emph{precise} bound $W_k^{j}$ (and hence $\rho_k^j$), due to Nelissen et al.\ \cite{ecrts15nelissen},  has exponential time complexity (even for the considered special case of a single self-suspending task). If over-approximations can be tolerated, finding a safe upper bound on $W_k^j$ for $j=1,2,\ldots,m_k-1$ can be done in pseudo-polynomial time \cite{PH:rtss98,Huang:multiseg}, albeit at a loss of accuracy and hence generality.
